\section{Multivariate relational data example}
\label{sec:NBA}

In this section, we demonstrate the applicability of our estimators to
multivariate relational data.  Such data may be viewed as a three-way
tensor $\mathcal{X}$ where entry $\mathcal{X}_{[i,j,k]}$ is the value
of relation type $k$ from node $i$ to node $j$. One example of such a
data set is a social network in which multiple types of relations are
measured between individuals.  As another example, in sports
statistics, round robin interaction data consist of outcomes of
competitions between teams. In this section we illustrate our methods
with round robin data from the 2014-2015 regular season of the
National Basketball Association (NBA). The NBA consists of a Western
conference and an Eastern conference of fifteen teams each, where
intra-conference play has three to four games per year per pair of
teams and inter-conference play is limited to two games a season per
pair of teams. For each conference, we created a four dimensional
tensor where element $\mathcal{Y}_{[i,j,k,\ell]}$ is statistic $k$
obtained by team $i$ while playing team $j$ either during team $i$'s
first home ($\ell = 1$) or first away ($\ell = 2$) game against team
$j$ during the season. The statistics we considered were free-throw
percentage, two-point field goal percentage, and three-point field
goal percentage. We thus have two tensors each of dimension $15 \times
15 \times 3 \times 2$, one for each of the two conferences.  In this
section, we illustrate the utility of tensor shrinkage by predicting
late season relational basketball statistics from early season
data. Our approach is analogous to that of \cite{efron1975data}, who
illustrated the utility of vector shrinkage estimation by predicting
late season baseball batting averages from data on early season
batting averages.

The statistics in our data set are all empirical proportions. We model
the elements of $\mathcal{Y}$ with a binomial model,
\begin{align*}
  n_{i,j,k,\ell}\mathcal{Y}_{[i,j,k,\ell]} \sim \Bin(n_{i,j,k,\ell},p_{i,j,k,\ell}),
\end{align*}
where all elements are independent, given the $p_{i,j,k,\ell}$'s. We
apply an arc-sin transformation to the data tensor to stabilize the
variance:
\begin{align*}
  \mathcal{X}_{[i,j,k,\ell]} = (n_{i,j,k,\ell})^{1/2}\arcsin(2\mathcal{Y}_{[i,j,k,\ell]} - 1).
\end{align*}
From the central limit theorem, we have approximately
\begin{align*}
  \mathcal{X}_{[i,j,k,\ell]} \sim N(\Theta_{[i,j,k,\ell]},1),
\end{align*}
where $\Theta_{[i,j,k,\ell]} =
(n_{i,j,k,\ell})^{1/2}\arcsin(2p_{i,j,k,\ell} - 1)$, resulting in the
model in (\ref{equation:normal.model}).

A commonly used representation of a mean tensor $\Theta$ is an ANOVA
decomposition, such as
\begin{align*}
  \Theta_{[i,j,k,\ell]} = \mu + \alpha_i + \beta_j + \gamma_k + \delta_{\ell} + \tilde{\Theta}_{[i,j,k,\ell]},
\end{align*}
where $\tilde{\Theta}_{[i,j,k,\ell]}$ contains all of the interaction
effects. Note that $\mathbf{1}_{p_1}^T\alpha = 0$,
$\mathbf{1}_{p_2}^T\beta = 0$, $\mathbf{1}_{p_3}\gamma = 0$, and
$\mathbf{1}_{p_4}^T\delta = 0$, where $\mathbf{1}_{p_k}$ is the vector
of ones of length $p_k$. The tensor $\tilde{\Theta}$ also satisfies
$\Tilde{\Theta}_{(k)}\mathbf{1}_{p/p_k} = 0$ for all $k =
1,2,3,4$. Suppose we obtain the maximum likelihood estimates of $\mu$,
$\alpha$, $\beta$, $\gamma$, and $\delta$ by fitting a main-effects
ANOVA model. We then calculate the residual tensor,
\begin{align*}
%%\mathcal{R}_{[i,j,k,\ell]} = \mathcal{X}_{[i,j,k,\ell]} - \mathcal{X}_{[i\cdot\cdot\cdot]} - \mathcal{X}_{[\cdot j\cdot\cdot]} - \mathcal{X}_{[\cdot\cdot k\cdot]} - \mathcal{X}_{[\cdot\cdot\cdot\ell]} + 3\mathcal{X}_{[\cdot\cdot\cdot\cdot]},
  \mathcal{R}_{[i,j,k,\ell]} =& \mathcal{X}_{[i,j,k,\ell]} -
  \frac{p_1}{p}\sum_{j',k',\ell'}\mathcal{X}_{[i,j',k',\ell']} -
  \frac{p_2}{p}\sum_{i',k',\ell'}\mathcal{X}_{[i',j,k',\ell']} -
  \frac{p_3}{p}\sum_{i',j',\ell'}\mathcal{X}_{[i',j',k,\ell']} \\
  &- \frac{p_4}{p}\sum_{i',j',k'}\mathcal{X}_{[i',j',k',\ell]} + \frac{3}{p}\sum_{i',j',k',\ell'}\mathcal{X}_{[i',j',k',\ell']}.
\end{align*}
This residual tensor has an expected value of $\tilde{\Theta}$.  It
was proposed in \cite{stein1966approach} and \cite{efron1972empirical}
that we estimate the interaction effects $\tilde{\Theta}$ with a
vector shrinkage-type estimator on the residuals. If the interactions
$\tilde{\Theta}$ are close to zero --- when the interaction effects
are small --- then such estimators will adaptively shrink the
residuals towards zero. However, these estimators were developed to
adapt to patterns in vectors or matrices of residuals, and not tensors
of residuals. In contrast, our approach should be able to adapt to
these patterns along any of the four modes of the residual tensor.

We applied mode-specific soft-thresholding and the truncated HOSVD to
the array of residuals $\mathcal{R}$ from the main effects ANOVA
model. These methods suggest that the residual tensor should be
heavily shrunk both towards zero and towards low multilinear rank
structure. For the West, the Frobenius norm of the residual tensor was
38.38, while the Frobenius norm of the resulting shrunken residual
tensor using the mode-specific soft-thresholding estimator was
7.81. In the East, the values were 38.95 and 6.97, respectively. We
also used SURE to estimate the multilinear rank of each residual
tensor using the truncated HOSVD. The estimated multilinear rank of
the residual tensor of the Western conference was $2 \times 3 \times 1
\times 2$, and for the Eastern conference the estimated multilinear
rank was $4 \times 2 \times 1 \times 1$. These are very small ranks
compared to the dimensions of the tensors $15 \times 15 \times 3
\times 2$.

An ad hoc evaluation of the performance of our estimators can be
obtained by predicting game statistics after the first home and first
away games. Since some teams only play each other three times, we do
not have late season data on all possible combinations of team pairs
by home versus away games. For the late season data we do have, we
present the squared error losses for predicting the statistics of the
remaining part of the season for each conference in Table
\ref{tab:sel.nba}. The different estimators are (1) the raw data array
$\mathcal{X}$, (2) the mean estimates of the main-effects ANOVA model,
(3) the mode-specific soft-thresholding shrunken residual tensor added
to the mean estimates of the main-effects ANOVA model, (4) the
truncated HOSVD shrunken residual tensor added to the mean estimates
of the main-effects ANOVA model, and (5) an estimator derived from
logistic regression using the main-effects of each mode. The losses
are with respect to the arc-sin transformed data. The poor performance
of $\mathcal{X}$ is unsurprising. The amount of shrinkage that our
estimators produce indicates that the fully saturated model is
over-fitting and that most of the information is contained in the
main-effects. However, our mode-specific soft-thresholding estimator
is also fitting the fully saturated model and it performs comparable
to the main-effects ANOVA model, even improving the predictions for
the Eastern conference.


\begin{table}[ht]
  \centering
  \begin{tabular}{rrr}
    Estimator & East & West\\
    \hline
    $\mathcal{X}$ & 2410 & 2476\\
    ANOVA & 1344 & 1364 \\
    Mode-specific Soft-thresholding & 1327 & 1385 \\
    Truncated HOSVD & 1391 & 1451\\
    Logistic Regression & 1481 & 1552\\
  \end{tabular}
  \caption{Squared error losses when predicting the statistics of the remaining games of the season.}
  \label{tab:sel.nba}
\end{table}

\begin{comment}
  \begin{itemize}
  \item Our estimators adaptively shrunk residuals very close to
    zero. For the West, the Frobenius norm of the residual array was
    38.38, and norm of the resulting shrunken residual array was
    7.81. In the East, the values were 38.95 and 6.97.
  \item Also, performance on the remaining season was comparable: The
    loss in the West was 1385 vs loss of 1364 and in the East it was
    1327 vs 1344. For $X$ the loss was 2200 in the West and 2410 for
    the East. Note that not taking into account the tensor structure
    (by using $\mathcal{X}$, does horribly).
  \item West: estimated rank of $2 \times 3 \times 1 \times 2$.
  \item East: estimated rank of $4 \times 2 \times 1 \times 1$.
  \item East Multiway EDA:
    \begin{itemize}
    \item Third mode first singular vector: (-0.94,0.08,0.34) (for FT,
      2FG, 3FG)
    \item Hence: core array relationship changes between FT and 3FG,
      and most of the effects of 2FG represented in additive model.
    \item Fourth mode first singular vector: (0.38,0.92) (for home and
      away games).
    \item Hence, effect more pronounced for away games vs home games
      (away games more variable?).
    \item Resulting Core array (represents relationship between first
      and second modes, changes sign for free throws vs 3pt field
      goals):
      \begin{table}[ht]
        \centering
        \begin{tabular}{rr}
          1 & -6 \\
          -6 & -1 \\
          -4 & -2 \\
          1 & 0 \\
        \end{tabular}
      \end{table}
    \item Biggest interactions occur between (1) $U_{1[,1]}$ and
      $U_{2[,2]}$, (2) $U_{1[,2]}$ and $U_{2[,1]}$, and to a lesser
      extend (3) $U_{1[,3]}$ and $U_{2[,1]}$.
    \end{itemize}
  \end{itemize}
\end{comment}